\LARGE \bf Mean Estimation from One-Bit Measurements

### 2020-12-04 proof-adaptive-optimality.tex
Fixed typos. Updated Fig. "optimal density".  Rewrote Conclusions
minor changes to proofs of Sec IV

\begin{proof-of-lemma}[\ref{lem:fisher_bound}]  
  When $f$ is the normal density function, this lemma follows from
  \cite[Thm.~3]{Barnes2018}. The proof below is based on a different techique
  than in \cite{Barnes2018}, and is valid for any log-concave symmetric
  density satisfying Assumption~\ref{assump:failure_rate}. \\

  The Fisher information of $B$ with respect to $\theta$ is given by
  \begin{align}
    I_\theta & =  \ex{ \left( \frac{d}{d\theta} \log P\left( B | \theta \right) \right)^2 |\theta } \nonumber \\
    & = \frac{ \left(\frac{d}{d\theta} P(B=1|\theta) \right)^2}{P(B=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(B=-1|\theta) \right)^2} {P(B=-1| \theta)} \nonumber \\
    & =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(B=1 | \theta) \left(1-P(B=1|\theta) \right)  }, \nonumber \\
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
  \end{align}
  where differentiation under the integral sign in $(a)$ is possible since
  $f$ is log-concave and so a.e.\ differentiable
  (cf.~\cite{Bertsekas73}),
  with a.e.\ derivative $f'(x)$. Regularity of
  the Lebesgue measure implies that for any $\epsilon>0$, there exists a
  finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that
  \begin{equation*}
    \int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon,
  \end{equation*}
  which implies that for any $\epsilon' > 0$, the set $A$ in
  \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of
  disjoint intervals without increasing $I_\theta$ by more than
  $\epsilon'$. It is therefore enough to proceed in the proof assuming that
  $A$ is of the form
  \begin{equation*}
    A = \cup_{j=1}^k (a_j,b_j),
  \end{equation*}
  with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
  \begin{align*}
    \mathbb P(B_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
    & = \sum_{j=1}^k \left( F \left(b_j-\theta\right) -  F \left(a_j-\theta\right)  \right),
  \end{align*}
  so we may rewrite Eq.~\eqref{eq:lem_fisher_bound_proof1} as
  \begin{align*}
    I_\theta & =   \frac { \left( \sum_{j=1}^{k} f \left(a_j-\theta \right) - f \left( b_j-\theta \right)  \right)^2 } 
    { \left( \sum_{j=1}^k F \left( b_j-\theta \right) - F \left( a_j-\theta \right)  \right) }  \nonumber \\
    & \times \frac {1} 
    {1- \left( \sum_{j=1}^k F \left(  b_j-\theta \right) - F \left( a_j-\theta \right)  \right) } 
  \end{align*}
  It follows from Lemma~\ref{lem:bound_intervals} that for any $\theta \in
  \R$ and any choice of the intervals' endpoints,
  \begin{equation*}
    I_\theta \le
    \max_{t \in \{a_j,b_j, j=1,\ldots,k\} } 4f(t)^2 \leq 4 f(0)^2, 
  \end{equation*}
  where the last transition is by Assumption~\ref{assump:failure_rate}. 
\end{proof-of-lemma}

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}.
To bound the Fisher information of any set of $n$ one-bit
messages with respect to $\theta$,
%% we first note that without loss of
%% each message $B_i$ can is of the form
%% \begin{equation}
%% B_i = \begin{cases}
%% X_i \in A_i & 1, \\
%% X_i \notin A_i & -1,
%% \end{cases} 
%% \end{equation}
%% where $A_i \subset \R$ is a Borel measurable set. 
%% %Indeed, any measurable function $(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = B^{-1}(1)$.
we consider the conditional distribution $P({B_1,\ldots,B_n|\theta})$ of
$(B_1,\ldots,B_n)$ given $\theta$. We have
\begin{align*}
  P\left( B_1,\ldots,B_n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B_1,\ldots,B_{i-1} \right),
  %% \label{eq:adpt_lower_bound_proof:1}
\end{align*}
where $P\left(B_i =1 | \theta, B_1,\ldots,B_{i-1} \right) = \mathbb P\left(
X_i \in A_i\right)$, when $B_i = \indic{X_i \in A_i}$ by independence,
so that the Fisher information of $B_1,\ldots,B_n$ with
respect to $\theta$ is
\begin{align}
I_\theta(B_1,\ldots,B_N) = \sum_{i=1}^n I_\theta (B_i|B_1,\ldots,B_{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B_{i-1},\ldots,B_1)$ is the Fisher information of the
distribution of $B_i$ given $B_1,\ldots,B_{i-1}$.
Lemma~\ref{lem:fisher_bound} implies that $I_\theta
(B_i|B_{i-1},\ldots,B_1) \leq 4f(0)^2$. The Van Trees inequality
\cite{van2004detection, gill1995applications} now implies
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B_1,\ldots,B_n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}
\end{align*}
as desired.

\subsection{An Isoperimetric Lemma}
\label{sec:bound_intervals_delta}

@@

  $\Delta_N$. (We may take the compactification of $\R$ as $[-\infty, \infty]$.)

>>>>

\begin{proof-of-lemma}[\ref{lem:fisher_bound}]  
  We first note that in the special case where $f$ is a normal density, Lemma~\ref{lem:fisher_bound} follows from \cite[Thm.~3]{Barnes2018}. The proof below, valid for any log-concave symmetric density satisfying Assumption~\ref{assump:failure_rate}, is based on a different techique than that of \cite{Barnes2018}. \\

  Write the Fisher information of $B$ with respect to $\theta$ as
  \begin{align}
    I_\theta & =  \ex{ \left( \frac{d}{d\theta} \log P\left( B | \theta \right) \right)^2 |\theta } \nonumber \\
    & = \frac{ \left(\frac{d}{d\theta} P(B=1|\theta) \right)^2}{P(B=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(B=-1|\theta) \right)^2} {P(B=-1| \theta)} \nonumber \\
    & =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(B=1 | \theta) \left(1-P(B=1|\theta) \right)  }, \nonumber \\
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
  \end{align}
  where differentiation under the integral sign in $(a)$ is possible since
  $f$ is log-concave hence a.e.\ differentiable
  (cf.~\cite{Bertsekas73}) with a.e.\ derivative $f'(x)$. Regularity of
  the Lebesgue measure implies that for any $\epsilon>0$, there exists a
  finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that
  \begin{equation*}
    \int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon.
  \end{equation*}
  It follows that for any $\epsilon' > 0$, the set $A$ in
  \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of
  disjoint intervals without increasing $I_\theta$ by more than
  $\epsilon'$. Consequently, we may proceed assuming that
  $A$ is of the form
  \begin{equation*}
    A = \cup_{j=1}^k (a_j,b_j),
  \end{equation*}
  with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
  \begin{align*}
    \mathbb P(B_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
    & = \sum_{j=1}^k \left( F \left(b_j-\theta\right) -  F \left(a_j-\theta\right)  \right),
  \end{align*}
  so we may rewrite Eq.~\eqref{eq:lem_fisher_bound_proof1} as
  \begin{align*}
    I_\theta & =   \frac { \left( \sum_{j=1}^{k} f \left(a_j-\theta \right) - f \left( b_j-\theta \right)  \right)^2 } 
    { \left( \sum_{j=1}^k F \left( b_j-\theta \right) - F \left( a_j-\theta \right)  \right) }  \nonumber \\
    & \times \frac {1} 
    {1- \left( \sum_{j=1}^k F \left(  b_j-\theta \right) - F \left( a_j-\theta \right)  \right) } 
  \end{align*}
  It follows from Lemma~\ref{lem:bound_intervals} that for any $\theta \in
  \R$ and any choice of the intervals' endpoints,
  \begin{equation*}
    I_\theta \le
    \max_{t \in \{a_j,b_j, j=1,\ldots,k\} } 4f(t)^2 \leq 4 f(0)^2, 
  \end{equation*}
  where the last transition is by Assumption~\ref{assump:failure_rate}. 
\end{proof-of-lemma}

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}.
To bound the Fisher information of any set of $n$ one-bit
messages with respect to $\theta$, we consider the conditional distribution $P({B_1,\ldots,B_n|\theta})$ of
$(B_1,\ldots,B_n)$ given $\theta$. We have
\begin{align*}
  P\left( B_1,\ldots,B_n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B_1,\ldots,B_{i-1} \right),
  %% \label{eq:adpt_lower_bound_proof:1}
\end{align*}
where $P\left(B_i =1 | \theta, B_1,\ldots,B_{i-1} \right) = \mathbb P\left(
X_i \in A_i\right)$, when $B_i = \indic{X_i \in A_i}$ by independence,
so that the Fisher information of $B_1,\ldots,B_n$ with
respect to $\theta$ is
\begin{align}
I_\theta(B_1,\ldots,B_N) = \sum_{i=1}^n I_\theta (B_i|B_1,\ldots,B_{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B_{i-1},\ldots,B_1)$ is the Fisher information of the
distribution of $B_i$ given $B_1,\ldots,B_{i-1}$.
Lemma~\ref{lem:fisher_bound} implies that $I_\theta
(B_i|B_{i-1},\ldots,B_1) \leq 4f(0)^2$. The Van Trees inequality
\cite{van2004detection, gill1995applications} now implies
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B_1,\ldots,B_n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}
\end{align*}
as desired.

\subsection{Isoperimetric Lemma}
\label{sec:bound_intervals_delta}

@@

  $\Delta_N$. %(We may take the compactification of $\R$ as $[-\infty, \infty]$.)
